Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-08-01 SDSS Journal Paper/Paper Draft.Rmd

<!-- In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. This direction will be limited but may be useful for the sake of ease on the town leaders. In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce a flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey. -->

@@

\section {Conclusions}

@@

@@

<!-- With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but along all types of dashboard design. We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research. -->

<!-- When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm. Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incorporated earlier or more often. As a result both groups become frustrated.   -->

>>>>

<!-- Parallel Coordinate Plots (PCPs) allowed the user to look at multidimensional variables, while the Guided Discovery Learning was used to make sure that users were informed of the information that was presented, it did not provide the town leaders any more ease in understanding our statistical graphic. Our new proposed direction will be limited but may be useful for the sake of ease on the town leaders and allow for the exploration into other pieces of data, by using a flower plot method. This method may not represent that data in the way that could be misleading of the whole picture of the town, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey. -->

<!-- While working with a variety of disciplines, EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. Because EDA can be hard to understand when the idea of EDA is only to prepare for modeling. However, EDA can be used in a data discovery, such as the work displayed in this article. Data Discovery can help with the quality of the data, in order to find new resources. -->

@@

\section{Conclusions}

@@

<!-- In conclusion, using a less known statistical graphic, PCPs, to the novice users was used to make sure that the users were informed of the multidimensional aspect of their information to help data driven decision making. As a result, we decided to look for easier ways to display data, such as boxplots, barplots, or flower plots. This direction will be limited but may be useful for the sake of ease on the town leaders. -->

@@

<!-- With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls by Stephen Few could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but across all types of dashboard design. We are encouraged to minimize an interface that will not be overwhelming the user, but in what order should these components should be used can be a struggle for our research. -->

<!-- When showing our applications versions to internal research members, it was difficult to collective incorporate necessary components and empower collaborators to help brainstorm. Once a design was presented, there were polarizing responses from no feedback to a dramatic change of the dashboard layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incorporated earlier or more often. As a result both groups become frustrated. As a result our finalizing motto will be to "test often and test again," this will help ease the feeling of exclusivity with our collaborators.    -->

### 2022-08-01 SDSS Journal Paper/Paper-Draft.tex

Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary. This approach created a way for the user to see their town information at the top of the page, this framework in the design allowed for the highlighted pieces of the town in generally. Once, the town components at the top of the page allowed the user to see the PCP after understanding their own status. The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.

While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP. The need to explain a visual means that the user is more likely to not use the plot effectively. Our hope was that the PCP would be a bit more ``self-explanatory'' than it was initially. The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.

Another hurdle and underlying tone from rural towns would be the fact that a larger town would like to change the framework of the uniqueness of the town. As a result, small towns don't like to see that their information being displayed publicly or compared to another city. This could a huge barrier for our research and work. To make sure that this does not become a problem, in the future work making sure that when allowing the user to compare towns in a visual graphic with a de-identified towns in the UI. For example, the town user selects similar population size, the each town that is populated will have a collection of ten similar towns using a label, such as ``Town 1'', ``Town 2'', etc. Along with the de-identifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected. This login system can then be used as a filter to setup the dashboard there forward. This will hopefully remove the lack of ability to be interactive in the purpose original design but give towns the comfort in feeling that their town data is being secure.

Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit. By suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.

Support for the dashboard application should have a help email for towns to ensure results and interpretation are clear. In this proposed idea, email communication will be connected to our research group, in which the help will not only have statisticians but also city planners that can help answer questions. Our integrated system will be able to help towns with data driven decision-making less of an invasion of privacy but as a collaborative workflow that will help their town succeed.

@@

In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.). In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.

Parallel Coordinate Plots (PCPs) allowed the user to look at multidimensional variables, while the Guided Discovery Learning was used to make sure that users were informed of the information that was presented, it did not provide the town leaders any more ease in understanding our statistical graphic. Our new proposed direction will be limited but may be useful for the sake of ease on the town leaders and allow for the exploration into other pieces of data, by using a flower plot method. This method may not represent that data in the way that could be misleading of the whole picture of the town, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

While working with a variety of disciplines, EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. Because EDA can be hard to understand when the idea of EDA is only to prepare for modeling. However, EDA can be used in a data discovery, such as the work displayed in this article. Data Discovery can help with the quality of the data, in order to find new resources.

@@

In conclusion, using a less known statistical graphic, PCPs, to the novice users was used to make sure that the users were informed of the multidimensional aspect of their information to help data driven decision making. As a result, we decided to look for easier ways to display data, such as boxplots, barplots, or flower plots. This direction will be limited but may be useful for the sake of ease on the town leaders.

Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'\,' design principles stemming from more controlled experiments hold up in real-world situations.

In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls by Stephen Few could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but across all types of dashboard design. We are encouraged to minimize an interface that will not be overwhelming the user, but in what order should these components should be used can be a struggle for our research.

When showing our applications versions to internal research members, it was difficult to collective incorporate necessary components and empower collaborators to help brainstorm. Once a design was presented, there were polarizing responses from no feedback to a dramatic change of the dashboard layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incorporated earlier or more often. As a result both groups become frustrated. As a result our finalizing motto will be to ``test often and test again,'' this will help ease the feeling of exclusivity with our collaborators.

>>>>

Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary. This approach created a way for the user to see their town information at the top of the page and to explore the PCP after reviewing their own town's essential statistics. The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot, providing an opportunity to explore the wider dataset from a base of familiar knowledge.
{\textcolor{red}{While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start. We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.}}

One factor that we have briefly considered and have seen hints of in our user feedback is that towns may not want to be compared negatively with other towns. While users have very definite ideas about which towns they would like to compare to, we can always mask the town names and move back to comparisons based on town size and other factors (for instance, whether or not a town is the county seat is a factor that is important outside of population). Using this approach, we would label each town as ``Town 1'', ``Town 2'', and so on, which would eliminate some of the fears about negative comparisons, but would also remove some of the novelty of the data dashboard for our users and would prevent users from drawing on their own outside knowledge about each of the comparison towns.

We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the app and teach towns how to use it. We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining quality of life amid shrinking populations.

@@

One avenue we will explore in future iterations of the dashboard is to incorporate other dashboards generated by different groups within this project. This will create a wider field of information to explore: for instance, some of the additional work will focus on the 99 towns featured in the Iowa Small Town Poll; this will allow us to showcase survey-based measures of quality of life alongside the more objective measurements assembled in the dataset discussed in this paper. While at least one tab of this omni-dashboard will still focus on wider EDA and discovery, we hope to incorporate other information as well to provide a more well-rounded data display encompassing most of the facets of this complex project.

We are also mindful of a distinction between ``eye candy'' and purpose-driven data visualization. While we have typically focused on the latter, there is certainly a place in our dashboard for the former as well. ``Eye candy'' visualization is intended to draw the viewer in and motivate them to explore; while these visualizations may not be particularly effective at communicating quantitative information, if they motivate the user to engage with the rest of the dashboard, they still serve a purpose. It is with this mindset that we intend to explore the use of flower plots - the artistic opportunities combined with the display of quantitative information (even in a form that isn't optimal for quantitative comparisons) may be useful to engage viewers before transitioning to more useful data visualizations intended to provide accurate quantitative comparisons.

EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept. One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project. Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard. While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.

@@

In this paper, we have documented the process of designing a dashboard for exploration and visualization of a large and complex data set assembled from many different sources. Our primary audience was leaders of small towns in Iowa, with a secondary audience of researchers in fields other than statistics collaborating on this project with us. Through the process of revising our dashboard, we found that the idea of guided discovery learning as implemented in our first version did not work as well as we had anticipated. It was more important to focus on allowing users to explore their questions about the dataset by facilitating user-driven comparisons and exploration, rather than attempting to anticipate user desires by providing comparison towns. In addition, we found that it would be more effective to draw users in with novel visual displays, as these seemed to attract more interest than providing known facts and an opportunity to explore outwards from an initial area of familiarity.

While it is hard to apply the findings from one fairly specific visualization project more widely, there is a lack of resources in this area that provide both design philosophies and actual analysis of user feedback in a qualitative sense. We have attempted to address this dearth of information by providing the design strategies, user feedback, and our planned and executed modifications, in the hopes that others facing the daunting challenge of designing a dashboard for EDA may learn something from our experiences.
